---
title: "Evolving Worldviews: Factors Influencing the Shifting Perspectives of American Adults on Abortion" 
subtitle: "An Analysis of US General Social Survey in 2006 and 2016"
author: 
- Shirley Chen
thanks: "Code and data in this report are available at: https://github.com/shirleychen003/abortion.git."
date: today
date-format: "long" 
format: pdf
abstract: Debates surrounding abortion access continue to generate a divide within the United States population. Surveys extracted from the U.S. General Social Survey’s datasets were utilized to explore the different reasons why women may obtain abortions and the demographic factors that may affect the public’s opinion on the topic. My investigation revealed that the general population is more likely to support abortion if the reason is health related, and that religious and the senior age group are more likely to criticize abortion legalization laws. Overall, my findings showcase the diverse range of factors that may affect one’s perspective on abortion access, including potential rationales for abortion and relevant demographic factors. 
toc: true
number-sections: true
bibliography: bibliography.bib
---

```{r}
#| include: false
### Workspace Setup ###
library(tidyverse)
library(tinytex)
library(janitor)
library(here)
library(knitr)
library(usethis)
library(gitcreds)
library(kableExtra)
library(gt)
library(ggpubr)
library(patchwork)
library(arrow)
library(readr)

#### Import data ####
cleaned_GSS_data <-
  read_csv("../data/cleaned_GSS_data.csv")

filtered_GSS_data <-
  read_csv("../data/filtered_GSS_data.csv")

#### Variables ####
years <- c("2006", "2016")
responses <- c("yes", "no")
custom_yes_no_colours <- c("#78c1a3", "#f38989")
custom_yes_no_colours_reversed <- c("#f38989", "#78c1a3")

religion_order <- c("Protestant", "Catholic", "Jewish", "None", "Other", "Buddhism", 
                    "Hinduism", "Other eastern religions", "Muslim/Islam", 
                    "Orthodox-Christian" , "Christian", "Native American", 
                    "Inter-nondenominational")
reasons <- c("Rape", "Endangered Health", "Low Income", "Any Reason")

#2006 data
data_2006 <- cleaned_GSS_data |>
  filter(year==2006)

#2016 data
data_2016 <- cleaned_GSS_data |>
  filter(year==2016)


data_2006$religion <- factor(data_2006$religion, levels = religion_order)
data_2016$religion <- factor(data_2016$religion, levels = religion_order)

rape_data <- cleaned_GSS_data[c('year', 'rape')]
low_income_data <- cleaned_GSS_data[c('year', 'low_income')]
endangered_health_data <- cleaned_GSS_data[c('year', 'endangered_health')]
any_reason_data <- cleaned_GSS_data[c('year', 'any_reason')]

total_participants <-cleaned_GSS_data |>
  summarise(count = n())


```

# Introduction

Around 25.1 million abortion happen every year, with nearly 8% of maternal deaths being attributed to unsafe abortions according to the World Health Organization (WHO) [@abortionlaw]. Discussion about abortion access continues to be a contentious and polarizing matter within many societies around the world. One side of the battle, which is commonly referred to as the “pro-life” perspective, advocates for restrictive laws on abortions, arguing for the rights of the unborn and declaring the procedure as a moral issue. On the other side, the “pro-choice” side advocates for reproductive rights, arguing for the fundamental right that humans have to make their own choices with their bodies, which includes obtaining an abortion. One of the earliest instances related to abortion laws in the United States dates back to 1847, when the American Medical Association (AMA) was formed [@ppafund]. During this time, state legislatures moved to ban abortions since AMA members had launched a criminalization campaign against abortions. In the decades following, all the states were required to restrict abortions by law, with special exceptions involving health reasons. Restrictive laws on abortion access were normalized, and the stigma around the procedure continued to grow. Since then, the fight for reproductive rights has been characterized by both challenges and developments, with the U.S, Supreme Court implementing the infamous Roe v. Wade law protecting the right to an abortion in 1973, but then overturning the law in 2022 [@ppafund].

Abortion rates have reported to be declining on both a national and regional level in the United States. In a 2008 study, it is estimated that considering current abortion rates, around 1 in 4 U.S. women of the reproductive age will have an abortion by age 45 [@georgiapastor]. Additionally, abortion rates vary on the state, as they were reportedly increasing in Georgia in the last decade, with the rate of abortion being 16.9 abortions per 1000 women aged 15-44 in 2017. Despite abortions being more common in Georgia, individuals face numerous barriers within policy, systems, and in their communities, restricting safe access to legal abortions. Especially after the overturning of Roe V. Wade in 2022, access to affordable abortion care became heavily restricted, and now exists  on paper exclusively for people of already marginalized and oppressed communities [@usabortionpolicy]. Regardless of how “valid” reasons or circumstances may be, many states continue to hold restrictive policies with very limited access to legal abortions. Disapproval of abortions manifests not solely through laws and policies, but through the American citizens’ perspectives as well, encompassed by a spectrum of cultural, religious, and personal values. 

In this paper, we will look at multifaceted reasons why women may want to obtain an abortion and the public’s opinion on if they should be legally able to. Sourced from the U.S. General Statistical Survey (GSS) from NORC at the University of Chicago, I extract the variables representing the reasons of rape, endangered health, low income, and any reason, and consolidate demographic factors such as sex, age, and religion to provide an extensive analysis. Overall, it becomes evident that when considering the immediate health status of the women (rape and/or endangered health), respondents are more likely to support an abortion. A significant number of respondents have less sympathy when considering the socioeconomic status of the women or personal reasons that may not be explicitly disclosed, potentially disregarding the complexity in circumstances and individual autonomy. Religious affiliation and the older age group seem to further have a significant effect on the respondents’ opinion, demonstrating a discouraging outlook on the issue. 

In [Data], the data source and methodology, the data cleaning processes, biases and limitations,  and key variables that are interacted with are presented. In [Results], the relationships between demographic information of participants and their responses are analyzed with supportive graphs and tables. Finally, in [Discussion], the significance and possible causes of our findings are explored within the factors of religious beliefs, socioeconomic status, sex and age.


# Data

## Source Data and Methodology
Based at the University of Chicago since 1972, the GSS is a project with the objective of monitoring and analyzing the intricacies of American society [@GSSabout]. The GSS Data Explorer makes it so that data retrieved from the project is a publicly available resource, accessible to various types of people, such as educators, policymakers, or researchers through the National Opinion Research Center (NORC). It has even been referenced in leading publications, such as the New York Times and the Associated Press. Due to the variety of subjects covered regarding American society, the GSS is one of the most frequently visited resources for information for social sciences.

@tbl-cleaneddemographic displays the demographic data that was obtained from the dataset.

```{r}
#| label: tbl-cleaneddemographic
#| echo: false
#| warning: false
#| tbl-cap: Demographic data obtained from the GSS Dataset

variable_data <- data.frame( 
  Variable = c( "age", "sex", "relig"),
  "New Name" = c( "age", "sex", "religion"), 
  "Description" = c(
    "Respondent's age", 
    "Respondent's sex",
    "Respondent's religious preference"), 
  "Example Response" = c("42", "female", "Protestant")
)

kable(variable_data, 
      col.names = c("Variable", 
                    "New Name",
                    "Description" ,
                    "Example Response"),
      booktabs = TRUE,
      linesep = ""
      )
```

@tbl-cleanedvariables displays the data relating to opinions on potential reasons for abortion that were obtained from the dataset. Respondents were initially given the prompt: "Please tell me whether or not you think it should be possible for a pregnant woman to obtain a legal abortion if...". Afterwords, they were instructed to circle one answer for the each of four statements that finished the prompt:"The woman wants it for any reason?", "If the family has a very low income and cannot afford any more children?", "If the woman's own health is seriously endangered by the pregnancy?", and "If she became pregnant as a result of rape?". 

Majority of the GSS data was collected through face-to-face interviews with the target population of adults (18+) residing in the United States. The standard national survey methods were practiced, such as hiring interviewers and training supervisors when needed. Interviewers were required to complete a practice interview supervised by evaluators at NORC [@GSScodebook].

However, starting in 2002, computer-assisted personal interviewing (CAPI) methods were introduced [@GSScodebook]. The use of manual edits and keypunching were eliminated, and training to learn how to use CAPI was included. The dataset used for this paper was retrieved from the GSS Data Explorer website [@GSSexplorer]. All the survey data used to measure the public opinion on different reasons for abortion were in the Core Module of GSS and tagged with Abortion and Split Ballots; the specific variable names extracted were abany, abpoor, abhlth, and abrape. For the years and demographic data, the specific variable names extracted were year (later filtered to just 2006 and 2016), age, relig, and sex.

### Strengths and Limitations

GSS data was collected through surveys that were conducted using both online and in person methods. Training on recording data in real time was administered to the interviewers who conducted the in person surveys, expanding the scope of data collection methods and enhancing versatility. Cost effectiveness and accessibility has also been promoted by the GSS, demonstrated through the integration of computer-assisted personal interviewing (CAPI) methods for the questionnaires. The GSS is a known, reliable source that encompasses a vast range of topics, such as socioeconomic status, political attitudes, life satisfaction, etc [@GSSabout]. Additionally, the GSS expanded their scope to non-English speakers in 2006, encouraging inclusion and representatives in their data. The surveys continue to be administered, ensuring a comprehensive understanding of societal opinions and trends over time. 
Although the GSS possesses many strengths, there are limitations to consider within the organization. The online surveys may have a learning curve for the older generations due to limited technological literacy, which may cause older generations to not participate in surveys at all. That being said, conducting in person surveys is not the most cost-friendly solution, as more time and resources are usually involved. Biases are also more likely to happen when faced with interviewers in person. 

### Biases to Consider

When conducting research, many biases can affect the accuracy of your results. Especially in a survey about such a controversial and sensitive topic, biases are bound to arise. For example, a common bias that arises is the social desirability bias, where respondents will provide a response that seems more socially acceptable, regardless of their true beliefs [@surveybias]. Respondents may either show significant support or strong disapproval for the topic depending on the environment they are surrounded in. This also relates to cultural bias, as the cultural/societal norms that individuals are accustomed to can greatly affect their responses. Since the responses were mainly measured with “yes” and “no” responses, acquiescence bias can arise, which is when respondents are prone to constantly agreeing to the prompt despite their true belief. In order to mitigate this bias, it’s important to vary questions and answers and use multiple choice questions along with scale questions [@surveybias]. In addition, it would be favourable to include response choices that have more variety than just simple “yes” or “no” responses. 

## Data Cleaning

The open source statistical programming language [@R-base] was used to clean and analyze the data, along with producing the graphs and tables. The main packages that supported this process included [@R-tidyverse], [@R-ggplot2], [@R-knitr], and [@R-janitor].

The cleaning process involved filtering the specific data variables used for our analysis from the downloaded GSS dataset, and renaming any variables with meaningful names. For example, rather than “abany” being the column name for “The woman wants an abortion for any reason?” , we renamed it to ‘any_reason’, as shown in @tbl-cleanedvariables. Further, the numerical values representing the participants’ responses for the abortion (eg. Yes - 1, No - 2) and religion (eg. Protestant - 1, Catholic - 2, etc.) were changed to the representative words. @tbl-cleanedvariables shows the old and new variable names for the abortion questions used in cleaning, the description of variables, and sample responses. Further, data from only years 2006 and 2016 were obtained. 

```{r}
#| label: tbl-cleanedvariables
#| echo: false
#| warning: false
#| tbl-cap: GSS Dataset

variable_data <- data.frame( 
  Variable = c( "abany", "abpoor", "abhlth", "abrape"),
  "New Name" = c( "any_reason", "low_income", "endangered_health", "rape"), 
  "Description of Reasons for Abortion" = c(
    "the woman wants an abortion for any reason", 
    "the family has a low income",
    "the woman's own health is endangered",
    "she became pregnant as a result of rape"), 
  "Example Response" = c("yes", "no", "yes", "no")
)

kable(variable_data, 
      col.names = c("Variable", 
                    "New Name",
                    "Description of Reasons for Abortion" ,
                    "Example Response"),
      booktabs = TRUE,
      linesep = ""
      )
```


## Data Terminology

The response choices for the questions regarding abortion and their respective code in brackets are as follows: Inapplicable (-100), No Answer (-99), Do Not Know/Cannot Choose (-98), Skipped on Web (-97), Yes (1), and No (2). The response choices for the religion question and their respective code in brackets are as follows: No Answer (-99), Do Not Know/Cannot Choose (-98), Skipped on Web (-97), Protestant (1), Catholic (2), Jewish (3), None (4), Other (5), Buddhism (6), Hinduism (7), Other eastern religions (8), Muslim/Islam (9), Orthodox-Christian (10), Christian (11), Native American (12), and Inter-nondenominational (13). For our graphs, we did not include the Inapplicable, No Answer, and Do Not Know/Cannot choose responses to focus on the discernible participant responses.


## Responses

@fig-rape displays the proportion of Yes and No responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if she became pregnant as a result of rape?"

```{r}
#| label: fig-rape
#| echo: false
#| warning: false
#| fig-cap: Proportion of Yes and No Responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if she became pregnant as a result of rape?"

# Rape Case Responses
rape_percentages <- rape_data |>
  count(year, rape) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

rape_percentages$rape <- factor(
  rape_percentages$rape, levels =
    responses)

ggplot(rape_percentages, aes(
  x = as.factor(year),
  y = percentage, 
  fill = rape)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Year",
       y = "Percentage",
       fill = "Responses") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), 
                     limits = c(0, 100)) + 
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal()
```

@fig-endangeredhealth displays the proportion of Yes and No responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman's own health is seriously endangered by the pregnancy?"

```{r}
#| label: fig-endangeredhealth
#| echo: false
#| warning: false
#| fig-cap: Proportion of Yes and No Responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman's own health is seriously endangered by the pregnancy?"

# Endangered Health Responses
endangered_health_data <- cleaned_GSS_data[c('year', 'endangered_health')]

endangered_health_percentages <- endangered_health_data |>
  count(year, endangered_health) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

endangered_health_percentages$endangered_health <- factor(
  endangered_health_percentages$endangered_health, levels =
    responses)

ggplot(endangered_health_percentages, aes(
  x = as.factor(year),
  y = percentage, 
  fill = endangered_health)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Year",
       y = "Percentage",
       fill = "Responses") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), 
                     limits = c(0, 100)) + 
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal()
```

@fig-lowincome displays the proportion of Yes and No responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the family has a very low income and cannot afford any more children?"

```{r}
#| label: fig-lowincome
#| echo: false
#| warning: false
#| fig-cap: Proportion of Yes and No Responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the family has a very low income and cannot afford any more children?"

# Endangered Health Responses
low_income_percentages <- low_income_data |>
  count(year, low_income) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

low_income_percentages$low_income <- factor(
  low_income_percentages$low_income, levels =
    responses)

ggplot(low_income_percentages, aes(
  x = as.factor(year),
  y = percentage, 
  fill = low_income)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Year",
       y = "Percentage",
       fill = "Responses") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), 
                     limits = c(0, 100)) + 
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal()
```

Lastly, @fig-anyreason displays the proportion of Yes and No responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason?"

```{r}
#| label: fig-anyreason
#| echo: false
#| warning: false
#| fig-cap: Proportion of Yes and No Responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason?"

# Endangered Health Responses
any_reason_percentages <- any_reason_data |>
  count(year, any_reason) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

any_reason_percentages$any_reason <- factor(
  any_reason_percentages$any_reason, levels =
    responses)

ggplot(any_reason_percentages, aes(
  x = as.factor(year),
  y = percentage, 
  fill = any_reason)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Year",
       y = "Percentage",
       fill = "Responses") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), 
                     limits = c(0, 100)) + 
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal()
```

# Results
## Respondent Demographics

@tbl-gender shows the count and percentage of male and female respondents for 2006 and 2016. Responses were filtered to be strictly male or female. For both years, the percentage of female participants were higher than the male participants, as the female participant percentages were always above 50% while the male participant percentages stayed close to the mid 40% range. 

```{r}
#| label: tbl-gender
#| echo: false
#| warning: false
#| tbl-cap: The Proportion of Females and Males Involved in the Survey in 2006 and 2016

gender_demographic_data <- cleaned_GSS_data |>
  group_by(year, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

kable(gender_demographic_data, 
      col.names = c("Year", "Sex", "Count", "Percentage"),
      booktabs = TRUE,
      linesep = ""
      )
```

@tbl-agegroups displays the number of respondents among the different age groups for the 2006 and 2016 surveys. The classified age groups are ‘18-24’, ‘25-34’, ‘45-54’, ‘55-64’, and ‘65+’. The 18-24 age group had the least amount of participants each year, while the '35-44' age group had the highest proportion in 2006 and the '65+' age group had the highest proportion in 2016.

```{r}
#| label: tbl-agegroups
#| echo: false
#| warning: false
#| tbl-cap: Respondent Count of Participants Grouped by their Specified Age Groups for 2006 and 2016

age_group_2006 <- cleaned_GSS_data |>
  filter(year == 2006) |>
  mutate(age_group_2006 = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_group_2006) |>
  pivot_wider(names_from = age_group_2006, values_from = n)

age_group_2016 <- cleaned_GSS_data |>
  filter(year == 2016) |>
  mutate(age_group_2016 = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_group_2016) |>
  pivot_wider(names_from = age_group_2016, values_from = n)

# Put the Year column back
age_group_2006 <- mutate(age_group_2006, Year = 2006)
age_group_2016 <- mutate(age_group_2016, Year = 2016)

combined_age_group_table <- bind_rows(age_group_2006,
                                      age_group_2016)

combined_age_group_table <- combined_age_group_table |>
  select(Year, everything())

kable(combined_age_group_table,
      col.names = c("Year", "18-24", "25-34", "35-44", "45-54", "55-64", "65+"),
      booktabs = TRUE,
      linesep = ""
      )

```

@tbl-age shows that the age of the participants ranged from 18 to 89, with the average respondent being 47 years old in 2006 and 50 in 2016. Additionally, the mode age have a significant difference between the two years, with a mode value of 36 in 2006 and 52 in 2016. 

```{r}
#| label: tbl-age
#| echo: false
#| warning: false
#| tbl-cap: Total Respondent Mean, Median, Mode, Min, and Max Age by Year (2006 and 2016)

age_2006_data <- data_2006 |>
  select('age') |>
   summarise(
    mean = round(mean(age, na.rm = TRUE)),
    mode = as.numeric(names(sort(table(age), decreasing = TRUE))[1]),
    min = min(age, na.rm = TRUE),
    max = round(max(age, na.rm = TRUE))
  )

age_2016_data <- data_2016 |>
  select('age') |>
   summarise(
    mean = round(mean(age, na.rm = TRUE)),
    mode = as.numeric(names(sort(table(age), decreasing = TRUE))[1]),
    min = min(age, na.rm = TRUE),
    max = round(max(age, na.rm = TRUE))
  )

combined_age_data <- bind_rows(age_2006_data, 
                               age_2016_data)
combined_age_data <- cbind(years, combined_age_data)


kable(combined_age_data, 
      col.names = c("Year",
                    "Mean", 
                    "Mode", 
                    "Min", 
                    "Max"),
      booktabs = TRUE,
      linesep = ""
      )

```

@tbl-religion displays the proportion of participants of each religion in the survey in 2006 and 2016. The religion with the highest proportion of responses for both years was "Protestant" with 52.46% in 2006 and 47.70% in 2016. The religion with the lowest proportion of responses for 2006 was "Other eastern religions" with a value of 0.11%, and there was an equal amount of respondents for "Other eastern religions" and "Inter-nondenominational" with a value of 0.24% in 2016.

@fig-allreligions graphically illustrates the proportions of the religious beliefs of the participants.

```{r}
#| label: tbl-religion
#| echo: false
#| warning: false
#| tbl-cap: Proportion of Respondents by Identified Religion in 2006 and 2016

religion2006_data <- data_2006 |>
  group_by(religion) |>
  summarise(count = n()) |>
  mutate(
    religion = factor(religion, levels = religion_order),
    percentage = round((count / sum(count)) * 100, 2)
  )
religion2016_data <- data_2016 |>
  group_by(religion) |>
  summarise(count = n()) |>
  mutate(
    religion = factor(religion, levels = religion_order),
    percentage = round((count / sum(count)) * 100, 2)
  )

religion2006_data_2 <- religion2006_data |>
  rename(count_2006 = count, percentage_2006 = percentage)

# Rename count and percentage columns in religion2016_cleaned_data
religion2016_data_2 <- religion2016_data |>
  rename(count_2016 = count, percentage_2016 = percentage)

# Combine the datasets using left_join
combined_data <- left_join(religion2006_data_2, 
                           religion2016_data_2,
                           by = "religion")

kable(combined_data,
      col.names = c("Religion",
      "2006 Count",
      "2006 Percentages",
      "2016 Count",
      "2016 Percentages"), 
      booktabs = TRUE,
      linesep = "")
```

```{r}
#| label: fig-allreligions
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Graph Visually Displaying the Number of Respondents grouped by Identified Religion

religion_data <- cleaned_GSS_data |>
  count(religion) |>
  group_by(religion)

religion_data$religion <- factor(
  religion_data$religion, levels =
    religion_order)

ggplot(religion_data, aes(x = religion, y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Religion", y = "Number of Respondents") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```
## Overall Trends

@tbl-avgresponses summarizes the average of the responses for each prompt, where 1 presents a "yes" response and 2 represents a "no" response. Each reason shows a decrease in the average from 2006 to 2016, showing that more participants are responding with "yes" in the later year. The most significant change was in the "Any Reason" reasoning, with a decrease of 0.06 in the value. Otherwise, the deceases remained minimal in the other prompts, ranging from 0.005 to 0.026. The decreases in value could be attributed to the change in societal attitudes and increased awareness of the issue at hand. 

```{r}
#| label: tbl-avgresponses
#| tbl-cap: The Average of Responses by Year grouped by Prompt Regarding Abortion
#| warning: false
#| echo: false

rape_avg <- filtered_GSS_data |>
  group_by(year) |>
  summarize(
    avg_rape = round(mean(abrape), 3))

endangered_health_avg <- filtered_GSS_data |>
  group_by(year) |>
  summarize(
    avg_endangered_health = round(mean(abhlth), 3))

low_income_avg <- filtered_GSS_data |>
  group_by(year) |>
  summarize(
    avg_low_income = round(mean(abpoor), 3))

any_reason_avg <- filtered_GSS_data |>
  group_by(year) |>
  summarize(
    avg_any_reason = round(mean(abany), 3))

combined_table <- cbind(
  rape_avg,
  endangered_health_avg[, -1],
  low_income_avg[, -1],
  any_reason_avg[, -1])

kable(combined_table,
      format = "markdown", 
      col.names = c("Year",
                    "Rape",
                    "Endangered Health",
                    "Low Income",
                    "Any Reason"),
      digits = 3,
      booktabs = TRUE,
      linesep = "")
```

@fig-allproportions2006 displays the proportion of "yes" and "no" responses among all the variables in 2006. The rape and endangered health, have a noticeably higher amount of respondents who answered yes, with both proportions being higher then 75%. However, for the other two reasons, low income and any reason, a lower proportion of respondents responded yes, as they were both less than 50%. The differences in proportions could convey the US population's additional sympathy for health reasons that are beyond the mother's control.

@fig-allproportions2016 displays the proportion of "yes" and "no" responses among all the variables in 2016, with the patterns being very similar to the 2006 data. Although minimal, there is an increase in "yes" responses for the low income and any reason variables, illustrating a possible shift in societal attitudes regarding the issue. 

```{r}
#| label: fig-allproportions2006
#| echo: false
#| message: false
#| warning: false
#| fig-cap: Percentages of Yes and No Responses to All Prompts Regarding Abortion in 2006

# Prepare data
variable_data_2006 <- data_2006[c('rape', 
                                'low_income',
                                'endangered_health',
                                'any_reason')]

# Rape
rape_data_2006 <- data_2006[c('year','rape')] 

rape_percentages <- rape_data_2006 |>
  count(year, rape) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

rape_percentages$rape <- factor(
  rape_percentages$rape, levels =
    responses)

rape_proportions <- 
  ggplot(rape_percentages, mapping = aes(x = " ", y = percentage, fill = rape)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Rape", y = "Proportion for each response (%)", fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Low Income
low_income_data_2006 <- data_2006[c('year','low_income')] 
low_income_percentages <- low_income_data_2006 |>
  count(year, low_income) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

low_income_percentages$low_income <- factor(
  low_income_percentages$low_income, levels =
    responses)

low_income_proportions <- 
  ggplot(low_income_percentages, mapping = aes(x = " ", y = percentage, fill = low_income)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Low Income", 
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Endangered Health
endangered_health_data_2006 <- data_2006[c('year','endangered_health')] 
endangered_health_percentages <- endangered_health_data_2006 |>
  count(year, endangered_health) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

endangered_health_percentages$endangered_health <- factor(
  endangered_health_percentages$endangered_health, levels =
    responses)

endangered_health_proportions <- 
  ggplot(endangered_health_percentages, mapping = aes(x = " ", 
                                                      y = percentage,
                                                      fill = endangered_health)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Endangered Health",
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Any Reason
any_reason_data_2006 <- data_2006[c('year','any_reason')] 
any_reason_percentages <- any_reason_data_2006 |>
  count(year, any_reason) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

any_reason_percentages$any_reason <- factor(
 any_reason_percentages$any_reason, levels =
    responses)

any_reason_proportions <-
  ggplot(any_reason_percentages , mapping = aes(x = " ",
                                                y = percentage, 
                                                fill = any_reason)) +
  geom_col() +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Any Reason",
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

any_reason_proportion <- ggplot(any_reason_percentages, aes(x = year, y = percentage, fill = any_reason)) +
  geom_col() +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Any Reason", y = "", fill = "Response") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

combined_graph_2006 <- rape_proportions + 
  low_income_proportions +
  endangered_health_proportions +
  any_reason_proportions +
  plot_layout(nrow = 1)

combined_graph_2006
```

```{r}
#| label: fig-allproportions2016
#| echo: false
#| message: false
#| warning: false
#| fig-cap: Percentages of Yes and No Responses to All Prompts Regarding Abortion in 2016
# Prepare data
variable_data_2016 <- data_2016[c('rape', 
                               'low_income',
                                'endangered_health',
                                'any_reason')]

# Rape
rape_data_2016 <- data_2016[c('year','rape')] 

rape_percentages <- rape_data_2016 |>
  count(year, rape) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

rape_percentages$rape <- factor(
  rape_percentages$rape, levels =
    responses)

rape_proportions <- 
  ggplot(rape_percentages, mapping = aes(x = " ", y = percentage, fill = rape)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Rape", y = "Proportion for each response (%)", fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Low Income
low_income_data_2016 <- data_2016[c('year','low_income')] 
low_income_percentages <- low_income_data_2016 |>
  count(year, low_income) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

low_income_percentages$low_income <- factor(
  low_income_percentages$low_income, levels =
    responses)

low_income_proportions <- 
  ggplot(low_income_percentages, mapping = aes(x = " ", y = percentage, fill = low_income)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Low Income", 
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Endangered Health
endangered_health_data_2016 <- data_2016[c('year','endangered_health')] 
endangered_health_percentages <- endangered_health_data_2016 |>
  count(year, endangered_health) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

endangered_health_percentages$endangered_health <- factor(
  endangered_health_percentages$endangered_health, levels =
    responses)

endangered_health_proportions <- 
  ggplot(endangered_health_percentages, mapping = aes(x = " ", 
                                                      y = percentage,
                                                      fill = endangered_health)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Endangered Health",
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Any Reason
any_reason_data_2016 <- data_2016[c('year','any_reason')] 
any_reason_percentages <- any_reason_data_2016 |>
  count(year, any_reason) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

any_reason_percentages$any_reason <- factor(
 any_reason_percentages$any_reason, levels =
    responses)

any_reason_proportions <-
  ggplot(any_reason_percentages , mapping = aes(x = " ",
                                                y = percentage, 
                                                fill = any_reason)) +
  geom_col() +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Any Reason",
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

any_reason_proportion <- ggplot(any_reason_percentages, aes(x = year, y = percentage, fill = any_reason)) +
  geom_col() +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Any Reason", y = "", fill = "Response") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

combined_graph_2016 <- rape_proportions + 
  low_income_proportions +
  endangered_health_proportions +
  any_reason_proportions +
  plot_layout(nrow = 1)

combined_graph_2016
```


### Age

Majority of the respondents were within the age range of 24-60 years old, as shown in @fig-allages. The highest point was at age 58 with 81 respondents and the lowest point of the graph was age 88 with 7 respondents. Despite the younger population showing more of an involvement in the topic, @fig-allages shows that there is still a substantial portion of the older generation contributing their opinion, conveying the importance of the topic among all ages. 

```{r}
#| label: fig-allages
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Number of Respondents by Age

age_data <- cleaned_GSS_data |>
  count(age) |>
  group_by(age)

ggplot(age_data, mapping = aes(x = age, y = n)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 25)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 25)) +
  labs(x = "Age", y = "Number of Respondents") +
  theme_classic() 

```

More specifically, the proportion of respondents organized by age group are shown in @tbl-age_rape, @tbl-age_endangeredhealth, 
@tbl-age_lowincome, and @tbl-age_anyreason, for the abortion reasons of rape, endangered health, low income, and any reason, respectively. The 65+ age group consistently demonstrated the highest proportion in no responses for each abortion reason, whereas the age group for yes responses varied, but stayed close in range. This may be correlated to the fact that older generations often hold more traditional and conservative views, in which abortion is not favoured upon. The 35-44 age group showed the highest proportion in yes responses for the reasons of rape, low income, and any reason, while the 45-54 age group held the highest amount of responses for the endangered health reason. The 35-54 age group may have more synpathy and put more emphasis on socioeconomic factors, health, and individuality compared to the other age groups. 

```{r}
#| label: tbl-age_rape
#| tbl-cap: Response Percentage Proportions to Question "Should it be possible for a pregnant woman to obtain a legal abortion if she became pregnant as a result of rape?" by Age
#| warning: false
#| echo: false
#| message: false
age_rape_data <- cleaned_GSS_data[c('age','year', 'rape')]

age_group_rape <- age_rape_data |>
  group_by(rape) |>
  mutate(age_rape_data = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_rape_data) |>
  pivot_wider(names_from = age_rape_data, values_from = n) |>
  rename("Response" = `rape`) 

age_group_rape_percentage <- age_group_rape |>
  mutate(`18-24` = round(`18-24` / total_participants$count * 100, 2),
         `25-34` = round(`25-34` / total_participants$count * 100, 2),
         `35-44` = round(`35-44` / total_participants$count * 100, 2),
         `45-54` = round(`45-54` / total_participants$count * 100, 2),
         `55-64` = round(`55-64` / total_participants$count * 100, 2),
         `65+` = round(`65+` / total_participants$count * 100, 2))

switched_rows <- rbind(age_group_rape_percentage[2, ], age_group_rape_percentage[1, ], age_group_rape_percentage[-c(1, 2), ])

kable(switched_rows, 
      booktabs = TRUE,
      linesep = "")

```

```{r}
#| label: tbl-age_endangeredhealth
#| tbl-cap: Response Percentage Proportions to Question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman's own health is seriously endangered by the pregnancy?" by Age
#| warning: false
#| echo: false
#| message: false
age_EH_data <- cleaned_GSS_data[c('age','year', 'endangered_health')]

age_group_EH <- age_EH_data |>
  group_by(endangered_health) |>
  mutate(age_EH_data = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_EH_data) |>
  pivot_wider(names_from = age_EH_data, values_from = n) |>
  rename("Response" = `endangered_health`) 

age_group_EH_percentage <- age_group_EH |>
  mutate(`18-24` = round(`18-24` / total_participants$count * 100, 2),
         `25-34` = round(`25-34` / total_participants$count * 100, 2),
         `35-44` = round(`35-44` / total_participants$count * 100, 2),
         `45-54` = round(`45-54` / total_participants$count * 100, 2),
         `55-64` = round(`55-64` / total_participants$count * 100, 2),
         `65+` = round(`65+` / total_participants$count * 100, 2))

switched_rows <- rbind(age_group_EH_percentage[2, ], age_group_EH_percentage[1, ], age_group_EH_percentage[-c(1, 2), ])

kable(switched_rows, 
      booktabs = TRUE,
      linesep = "")

```

```{r}
#| label: tbl-age_lowincome
#| tbl-cap: Response Percentage Proportions to Question "Should it be possible for a pregnant woman to obtain a legal abortion if the family has a very low income and cannot afford any more children?" by Age
#| warning: false
#| echo: false
#| message: false
age_LI_data <- cleaned_GSS_data[c('age','year', 'low_income')]

age_group_LI <- age_LI_data |>
  group_by(low_income) |>
  mutate(age_LI_data = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_LI_data) |>
  pivot_wider(names_from = age_LI_data, values_from = n) |>
  rename("Response" = `low_income`) 

age_group_LI_percentage <- age_group_LI |>
  mutate(`18-24` = round(`18-24` / total_participants$count * 100, 2),
         `25-34` = round(`25-34` / total_participants$count * 100, 2),
         `35-44` = round(`35-44` / total_participants$count * 100, 2),
         `45-54` = round(`45-54` / total_participants$count * 100, 2),
         `55-64` = round(`55-64` / total_participants$count * 100, 2),
         `65+` = round(`65+` / total_participants$count * 100, 2))

switched_rows <- rbind(age_group_LI_percentage[2, ], age_group_LI_percentage[1, ], age_group_LI_percentage[-c(1, 2), ])

kable(switched_rows, 
      booktabs = TRUE,
      linesep = "")

```

```{r}
#| label: tbl-age_anyreason
#| tbl-cap: Response Percentage Proportions to Question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason?" by Age
#| warning: false
#| echo: false
#| message: false
age_AR_data <- cleaned_GSS_data[c('age','year', 'any_reason')]

age_group_AR <- age_AR_data |>
  group_by(any_reason) |>
  mutate(age_AR_data = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_AR_data) |>
  pivot_wider(names_from = age_AR_data, values_from = n) |>
  rename("Response" = `any_reason`) 

age_group_AR_percentage <- age_group_AR |>
  mutate(`18-24` = round(`18-24` / total_participants$count * 100, 2),
         `25-34` = round(`25-34` / total_participants$count * 100, 2),
         `35-44` = round(`35-44` / total_participants$count * 100, 2),
         `45-54` = round(`45-54` / total_participants$count * 100, 2),
         `55-64` = round(`55-64` / total_participants$count * 100, 2),
         `65+` = round(`65+` / total_participants$count * 100, 2))

switched_rows <- rbind(age_group_AR_percentage[2, ], age_group_AR_percentage[1, ], age_group_AR_percentage[-c(1, 2), ])

kable(switched_rows, 
      booktabs = TRUE,
      linesep = "")

```

### Sex

The gender demographic along with the responses for each abortion reason are exhibited and filtered by the years 2006 and 2016. The percentages are grouped by the response type of the participant, and then further grouped into their gender.
For each table, females made up a higher percentage for both responses in both of the years. Females showed the most support for abortions for the reason of endangered health (@tbl-endangeredhealthdemographic) and the least amount of support for abortions for any reason (@tbl-anyreasondemographic) in 2006 and  low income (@tbl-lowincomedemographic) in 2016.
Males also showed the most support for abortions in the case where the mother's health is endangered and showed the least amount of support for abortions that are obtained for any reason. While both genders conveyed a lot of support for abortions when there is risk of endangered health, participants also showed support for abortions for victims who were impregnated via rape (@tbl-rapedemographic).

```{r}
#| label: tbl-rapedemographic
#| tbl-cap: Rape Response Count and Percentage Proportions grouped by Sex
#| warning: false
#| echo: false

# 2006 table
rape_demographic_2006 <- data_2006[c('rape','sex')] 

rape_demographic_2006 <- rape_demographic_2006 |>
  mutate(response = factor(rape, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# 2016 table
rape_demographic_2016 <- data_2016[c('rape','sex')] 

rape_demographic_2016 <- rape_demographic_2016 |>
  mutate(response = factor(rape, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# Combine tables
combined_R_table <- inner_join(rape_demographic_2006,
                               rape_demographic_2016, 
                               by = c("response", "sex"))
kable(combined_R_table,
      col.names = c("Response",
                    "Sex",
                    "Count",
                    "Percentage",
                    "Count",
                    "Percentage"),
      booktabs = TRUE,
      linesep = "") |>
  add_header_above(c( " " = 2, "2006" = 2, "2016" = 2), bold = TRUE)
```

```{r}
#| label: tbl-endangeredhealthdemographic
#| tbl-cap: Endangered Health Response Count and Percentage Proportions grouped by Sex
#| warning: false
#| echo: false

# 2006 table
endangered_health_demographic_2006 <-
  data_2006[c('endangered_health','sex')] 

endangered_health_demographic_2006 <-
  endangered_health_demographic_2006 |>
  mutate(response = factor(endangered_health, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# 2016 table
endangered_health_demographic_2016 <- 
  data_2016[c('endangered_health','sex')] 

endangered_health_demographic_2016 <-
  endangered_health_demographic_2016 |>
  mutate(response = factor(endangered_health, 
                           levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# Combine tables
combined_EH_table <- inner_join(endangered_health_demographic_2006,
                               endangered_health_demographic_2016, 
                               by = c("response", "sex"))
kable(combined_EH_table,
      col.names = c("Response",
                    "Sex",
                    "Count",
                    "Percentage",
                    "Count",
                    "Percentage"),
      booktabs = TRUE,
      linesep = "") |>
  add_header_above(c( " " = 2, "2006" = 2, "2016" = 2), bold = TRUE)
```

```{r}
#| label: tbl-lowincomedemographic
#| tbl-cap: Low Income Response Count and Percentage Proportions grouped by Sex
#| warning: false
#| echo: false

# 2006 table
low_income_demographic_2006 <-
  data_2006[c('low_income','sex')] 

low_income_demographic_2006 <- low_income_demographic_2006 |>
  mutate(response = factor(low_income, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# 2016 table
low_income_demographic_2016 <- data_2016[c('low_income','sex')] 

low_income_demographic_2016 <- low_income_demographic_2016 |>
  mutate(response = factor(low_income, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# Combine tables
combined_LI_table <- inner_join(low_income_demographic_2006,
                             low_income_demographic_2016, 
                             by = c("response", "sex"))
kable(combined_LI_table,
      col.names = c("Response",
                    "Sex",
                    "Count",
                    "Percentage",
                    "Count",
                    "Percentage"),
      booktabs = TRUE,
      linesep = "") |>
  add_header_above(c( " " = 2, "2006" = 2, "2016" = 2), bold = TRUE)
```

```{r}
#| label: tbl-anyreasondemographic
#| tbl-cap: Any Reason Response Count and Percentage Proportions grouped by Sex
#| warning: false
#| echo: false

# 2006 table
any_reason_demographic_2006 <- data_2006[c('any_reason','sex')] 

any_reason_demographic_2006 <- any_reason_demographic_2006 |>
  mutate(response = factor(any_reason, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# 2016 table
any_reason_demographic_2016 <- data_2016[c('any_reason','sex')] 

any_reason_demographic_2016 <- any_reason_demographic_2016 |>
  mutate(response = factor(any_reason, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# Combine tables
combined_AR_table <- inner_join(any_reason_demographic_2006,
                               any_reason_demographic_2016, 
                               by = c("response", "sex"))
kable(combined_AR_table,
      col.names = c("Response",
                    "Sex",
                    "Count",
                    "Percentage",
                    "Count",
                    "Percentage"),
      booktabs = TRUE,
      linesep = "") |>
  add_header_above(c( " " = 2, "2006" = 2, "2016" = 2), bold = TRUE)
```

### Religious Views

Narrowing down the religions, @fig-religionprop displays the proportion of yes and no responses specifically for the Protestant and Non-Religious Respondents. The Protestant group was chosen since it was the religious group with the highest proportion of respondents, and the Non-Religious group was chosen to compare how religion may affect one's perspective on the issue. 
As we can see from @fig-religionprop, a higher proportion of protestant answered "no" consistently for each reason for abortion. These responses may be due to moral teachings and inteprations of life within the Christian religion, prompting more people to believe that abortion is taking away a life. Both Non-Religious and Protestant respondents showed the least support for rape cases or low income for reasons to obtain a legal abortion. Although the fraction of Protestants choosing "no" is higher, the graph illustrates how there is a shared perspective among the two groups regarding perspectives on abortion in cases involving rape or financial hardship. 

```{r}
#| label: fig-religionprop
#| fig-cap: Proportion of Responses for each Prompt filtered by the Protestant and Non-Religious Participants 
#| warning: false
#| echo: false

prot_none <- cleaned_GSS_data |>
  filter(religion %in% c("Protestant", "None")) |>
  select(religion, rape, low_income, endangered_health, any_reason)

prot_none_data <- prot_none |>
  pivot_longer(cols = c(rape, low_income, endangered_health, any_reason),
               names_to = "reason",
               values_to = "response")

ggplot(prot_none_data, aes(x = reason, fill = response)) +
  geom_bar(position = "fill", stat = "count") +
  facet_wrap(~ religion, ncol = 1) +
  labs(x = "Reasons for Abortion",
       y = "Proportion",
       fill = "Response") +
  scale_x_discrete(labels = reasons) +
  scale_fill_manual(values = custom_yes_no_colours_reversed) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  theme(legend.position = "right",
        axis.text.x = element_text(size = 8), 
        axis.title = element_text(size = 10),   
        strip.text = element_text(size = 14)) 
```

# Discussion

## Generational Differences
Generational differences are a contributing factor to how one may perceive obtaining an abortion. Gallup Analysis, an international public opinion data platform, indicated that senior American citizens aged 65 and older were the least supportive in legal abortions under any circumstances from 2005 to 2009. All the other age groups showed similar support for broadly legal abortion[@generationone] . However, young adults in the age range of 18 to 29 years old were more likely than the senior age group to say that abortion should be illegal in all circumstances. Access to diverse sources of education and information may explain these contrasting perspectives, as younger adults have had more access, generating more nuanced perspectives on the issue. 
Furthermore, the Public Religion Research Institute (PPRI) found that younger Americans or “Gen-Z” are more likely to support abortion legality than older Americans [@genzprochoice]. 72% of Americans within the Gen-Z and young Millennial generation believed that abortions should be legally attainable in all or majority of the cases, in comparison to the 64% of Americans who are aged 30 years old and older. 

One explanation may be the evolution of values and perspectives of women’s rights and freedom over time. Due to deep rooted misogyny and patriarchal structures within society, women were often not given many choices in their lives, including regarding reproductive health discussions. However, since younger generations have been more involved in feminist movements and advocate more for women’s rights, younger generations are more likely to support the women’s choice in abortion processes. 


## Sex
Sex plays a role in an individual’s perspective on abortions. The PPRI has discovered that young women and men differ in their attitudes about some of the restrictive abortion policies. Gen-Z women are more likely than men to express strong opposition against banning the use of FDA-approved abortion pills [@genzprochoice]. 52% of female Gen Z respondents strongly oppose the criminalization of abortion drugs, in contrast to 43% of male Gen Z respondents. Women often have a stronger opinion about abortion issues, since they are the ones physically, directly impacted by pregnancies and the potential consequences. Restrictions on abortions may also be recognized as a violation to a women’s sense of autonomy and bodily integrity. In various interviews conducted by the New York Times, men have expressed their difficulties to personalize with the issue, since they are unable to experience abortions and pregnancies [@gendergap]. Most of their opinions depend on the scenario that the pregnant woman is in, but that in itself also presents a grey area, as it can be a struggle to establish appropriate time constraints for legal abortion. Lastly, many men conveyed that abortion is not just an issue with women’s health, but with broader issues in society. Art Menichollo stated that the recent law implementations “try to take someone’s rights away”, conveying his support for legal abortions [@gendergap]. Overall, the perspectives of men widely differ, but due to their inability to empathize and experience a woman’s pregnancy for themselves, their comprehension of the complexities and nuances within reproductive rights are limited.

## Religious Views
Majority of the American population are religious, and most of the major religions contain teachings that disapprove of abortion. Because of this, researchers have found that there is a strong correlation between individual religiosity and negative abortion attitudes. In a polling, over 50% of respondents who held a religious affiliation with Protestantism or Catholicism stated that abortions were morally unacceptable, comparable to 20% of the survey population who held this belief but did not identify with any religious affiliation [@abortionmorality]. Additionally, in a survey with 5,387 abortion patients at a U.S. clinic, 28% of women were not spiritually at peace with their  abortion and 36% of women reported having spiritual concerns about their decision. With these statistics in mind, it is evident that the issue of abortion has been intertwined into religious beliefs, due to the complex, ethical and spiritual dimensions that arise. 

It is also important to consider the differing perspectives within each religion. After conducting interviews with numerous Protestant leaders, it was shown that the attitudes ranged  due to the complexity of the issue when connected to scripture. There were many grey areas to consider, such as the belief of when life officially begins and specific circumstances where abortions could be morally acceptable [@georgiapastor]. Due to the common rule in Christianity where compassion and empathy is to be expressed for your neighbours, many pastors stated they would ensure that they would support the women through their process. However, key similarities in pastoral care included recognizing the importance of life and supporting their advice with scripture, and informing women of the spiritual and emotional toll the abortion may take on them. Otherwise, pastors receive little to no training on these types of matters, since the majority of Christians often are not in favour of the abortion process. 

## Economic Considerations

The financial status of a pregnant woman and any future economic considerations are pivotal in determining if an abortion will be attained. The United States contains some of the highest pregnancy, childbirth, and postpartum costs on a global scale, with vaginal delivery costing around $ 11,453 USD on average and cesarean sections averaging $17,103 [@babycosts]. These costs are taking into account that there are no further medical complications that require further finances, or any pregnancy or postpartum care expenses. Having multiple babies or factors such as preexisting health conditions could potentially increase the expenses involved with the process. Additionally, although insurance considerably reduces the costs, it is estimated that $3,000 is required out of pocket for pregnancy and birth expenses. Without insurance, mothers are often required to pay $35,500 or more to healthcare providers [@babycosts]. If the mother were providing financial assistance to the child until they become an adult at age 18, the U.S. Agriculture Department has stated that they are expected to spend approximately $310,605 total, or around $17,000 a year [@babycosts]. Given all of this information, finances will also vary depending on the state, as there are different protection laws and costs by the state. 

Considering the federal minimum wage ($7.75 USD an hour) illustrates the difficulties that may arise when low-income families are trying to meet basic needs, along with providing sufficient childcare services. According to the Economic Policy Institute, a minimum wage worker must work 1,162 hours to cover annual costs for a 4 year old and 1,548 hours for an infant, which consists of working full time from January to July and January to September respectively [@childcosts]. With minimum wage increases not always mirroring the rising rate of inflation, financially unstable families often have a hard time making ends meet. This situation demonstrates the perpetuating poverty cycle within America, as although individuals are contributing to the economy by having children, the quality of life is not guaranteed due to the high costs of essential life services. The unpredictable and high costs that come with pregnancy and childcare are a major factor in a woman’s decision to obtain an abortion. In unexpected pregnancies, women of low economic status are often faced with the choice of continuing the pregnancy and needing to pay maintenance expenses, or terminating the pregnancy to avoid further financial and logistical complications. 

\newpage

# References
