---
title: "How Opinions On Abortion Have Evolved Among American Adults"
subtitle: "Analysis of US General Social Survey in 2006 and 2016"
author: Shirley Chen
thanks: "Code and data in this report are available at: https://github.com/shirleychen003/abortion.git."
date: today
date-format: "long" 
format: pdf
abstract: abstract in progress :)
toc: true
number-sections: true
bibliography: bibliography.bib
---

```{r}
#| include: false
### Workspace Setup ###
library(tidyverse)
library(tinytex)
library(janitor)
library(here)
library(knitr)
library(usethis)
library(gitcreds)
library(kableExtra)
library(gt)
library(ggpubr)
library(patchwork)
library(arrow)

#### Import data ####
cleaned_GSS_data <-
  read_csv("../data/cleaned_GSS_data.csv")

filtered_GSS_data <-
  read_csv("../data/filtered_GSS_data.csv")

#### Variables ####
years <- c("2006", "2016")
responses <- c("yes", "no")
custom_yes_no_colours <- c("#78c1a3", "#f38989")
custom_yes_no_colours_reversed <- c("#f38989", "#78c1a3")

religion_order <- c("Protestant", "Catholic", "Jewish", "None", "Other", "Buddhism", 
                    "Hinduism", "Other eastern religions", "Muslim/Islam", 
                    "Orthodox-Christian" , "Christian", "Native American", 
                    "Inter-nondenominational")
reasons <- c("Rape", "Endangered Health", "Low Income", "Any Reason")

#2006 data
data_2006 <- cleaned_GSS_data |>
  filter(year==2006)

#2016 data
data_2016 <- cleaned_GSS_data |>
  filter(year==2016)


rape_data <- cleaned_GSS_data[c('year', 'rape')]
low_income_data <- cleaned_GSS_data[c('year', 'low_income')]
endangered_health_data <- cleaned_GSS_data[c('year', 'endangered_health')]
any_reason_data <- cleaned_GSS_data[c('year', 'any_reason')]

total_participants <-cleaned_GSS_data |>
  summarise(count = n())


```

# Introduction

Discussion about abortion access continues to be a contentious and polarizing matter within many societies around the world. One side of the battle, which is commonly referred to as the “pro-life” perspective, advocates for restrictive laws on abortions, arguing for the rights of the unborn and declaring the procedure as a moral issue. On the other side, the “pro-choice” side advocates for reproductive rights, arguing for the fundamental right that humans have to make their own choices with their bodies, which includes obtaining an abortion. One of the earliest instances related to abortion laws in the United States dates back to 1847, when the American Medical Association (AMA) was formed (1). During this time, state legislatures moved to ban abortions since AMA members had launched a criminalization campaign against abortions. In the decades following, all the states were required to restrict abortions by law, with special exceptions involving health reasons. Restrictive laws on abortion access were normalized, and the stigma around the procedure continued to grow. Since then, the fight for reproductive rights has been characterized by both challenges and developments, with the U.S, Supreme Court implementing the infamous Roe v. Wade law protecting the right to an abortion in 1973, but then overturning the law in 2022 (1). 

In this paper, we will look at multifaceted reasons why women may want to obtain an abortion and the public’s opinion on if they should be legally able to. Sourced from the U.S. General Statistical Survey (GSS) from NORC at the University of Chicago, I extract the variables representing the reasons of rape, endangered health, low income, and any reason, and consolidate demographic factors such as sex, age, and religion to provide an extensive analysis. Overall, it becomes evident that when considering the immediate health status of the women (rape and/or endangered health), respondents are more likely to support an abortion. A significant number of respondents have less sympathy when considering the socioeconomic status of the women or personal reasons that may not be explicitly disclosed, potentially disregarding the complexity in circumstances and individual autonomy. 

In \[Data\], the data source and methodology, the data cleaning processes, biases and limitations,  and key variables that are interacted with are presented. In \[Results\], the relationships between demographic information of participants and their responses are analyzed with supportive graphs and tables. Finally, in \[Discussion\], the significance and possible causes of our findings are explored within the factors of religious beliefs, socioeconomic status, sex and age.


# Data

## Source Data and Methodology
Based at the University of Chicago since 1972, the GSS is a project with the objective of monitoring and analyzing the intricacies of American society [@GSSabout]. The GSS Data Explorer makes it so that data retrieved from the project is a publicly available resource, accessible to various types of people, such as educators, policymakers, or researchers through the National Opinion Research Center (NORC). It has even been referenced in leading publications, such as the New York Times and the Associated Press. Due to the variety of subjects covered regarding American society, the GSS is one of the most frequently visited resources for information for social sciences.

@tbl-cleaneddemographic displays the demographic data that was obtained from the dataset.

```{r}
#| label: tbl-cleaneddemographic
#| echo: false
#| warning: false
#| tbl-cap: Demographic data obtained from the GSS Dataset

variable_data <- data.frame( 
  Variable = c( "age", "sex", "relig"),
  "New Name" = c( "age", "sex", "religion"), 
  "Description" = c(
    "Respondent's age", 
    "Respondent's sex",
    "Respondent's religious preference"), 
  "Example Response" = c("42", "female", "Protestant")
)

kable(variable_data, 
      col.names = c("Variable", 
                    "New Name",
                    "Description" ,
                    "Example Response"),
      booktabs = TRUE,
      linesep = ""
      )
```

@tbl-cleanedvariables displays the data relating to opinions on potential reasons for abortion that were obtained from the dataset. Respondents were initially given the prompt: "Please tell me whether or not you think it should be possible for a pregnant woman to obtain a legal abortion if...". Afterwords, they were instructed to circle one answer for the each of four statements that finished the prompt:"The woman wants it for any reason?", "If the family has a very low income and cannot afford any more children?", "If the woman's own health is seriously endangered by the pregnancy?", and "If she became pregnant as a result of rape?". 

Majority of the GSS data was collected through face-to-face interviews with the target population of adults (18+) residing in the United States. The standard national survey methods were practiced, such as hiring interviewers and training supervisors when needed. Interviewers were required to complete a practice interview supervised by evaluators at NORC [@GSScodebook].

However, starting in 2002, computer-assisted personal interviewing (CAPI) methods were introduced [@GSScodebook]. The use of manual edits and keypunching were eliminated, and training to learn how to use CAPI was included. The dataset used for this paper was retrieved from the GSS Data Explorer website [@GSSexplorer]. All the survey data used to measure the public opinion on different reasons for abortion were in the Core Module of GSS and tagged with Abortion and Split Ballots; the specific variable names extracted were abany, abpoor, abhlth, and abrape. For the years and demographic data, the specific variable names extracted were year (later filtered to just 2006 and 2016), age, relig, and sex.

### Strengths and Limitations

### Biases to Consider

## Data Cleaning

The open source statistical programming language [@R-base] was used to clean and analyze the data, along with producing the graphs. The main packages that supported this process included [@R-tidyverse], [@R-ggplot2], [@R-knitr], [@R-janitor], and [@R-lubridate].

The cleaning process involved filtering the specific data variables used for our analysis from the downloaded GSS dataset, and renaming any variables with meaningful names. For example, rather than “abany” being the column name for “The woman wants an abortion for any reason?” , we renamed it to ‘any_reason’, as shown in @tbl-cleanedvariables. Further, the numerical values representing the participants’ responses for the abortion (eg. Yes - 1, No - 2) and religion (eg. Protestant - 1, Catholic - 2, etc.) were changed to the representative words. @tbl-cleanedvariables shows the old and new variable names for the abortion questions used in cleaning, the description of variables, and sample responses. Further, data from only years 2006 and 2016 were obtained. 

```{r}
#| label: tbl-cleanedvariables
#| echo: false
#| warning: false
#| tbl-cap: GSS Dataset

variable_data <- data.frame( 
  Variable = c( "abany", "abpoor", "abhlth", "abrape"),
  "New Name" = c( "any_reason", "low_income", "endangered_health", "rape"), 
  "Description of Reasons for Abortion" = c(
    "the woman wants an abortion for any reason", 
    "the family has a low income",
    "the woman's own health is endangered",
    "she became pregnant as a result of rape"), 
  "Example Response" = c("yes", "no", "yes", "no")
)

kable(variable_data, 
      col.names = c("Variable", 
                    "New Name",
                    "Description of Reasons for Abortion" ,
                    "Example Response"),
      booktabs = TRUE,
      linesep = ""
      )
```


## Data Terminology

The response choices for the questions regarding abortion and their respective code in brackets are as follows: Inapplicable (-100), No Answer (-99), Do Not Know/Cannot Choose (-98), Skipped on Web (-97), Yes (1), and No (2). The response choices for the religion question and their respective code in brackets are as follows: No Answer (-99), Do Not Know/Cannot Choose (-98), Skipped on Web (-97), Protestant (1), Catholic (2), Jewish (3), None (4), Other (5), Buddhism (6), Hinduism (7), Other eastern religions (8), Muslim/Islam (9), Orthodox-Christian (10), Christian (11), Native American (12), and Inter-nondenominational (13). For our graphs, we did not include the Inapplicable, No Answer, and Do Not Know/Cannot choose responses to focus on the discernible participant responses.


## Responses

### Rape

@fig-rape displays the proportion of Yes and No responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if she became pregnant as a result of rape?"

```{r}
#| label: fig-rape
#| echo: false
#| warning: false
#| fig-cap: Proportion of Yes and No Responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if she became pregnant as a result of rape?"

# Rape Case Responses
rape_percentages <- rape_data |>
  count(year, rape) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

rape_percentages$rape <- factor(
  rape_percentages$rape, levels =
    responses)

ggplot(rape_percentages, aes(
  x = as.factor(year),
  y = percentage, 
  fill = rape)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Year",
       y = "Percentage",
       fill = "Responses") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), 
                     limits = c(0, 100)) + 
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal()
```

### Endangered Health

@fig-endangeredhealth displays the proportion of Yes and No responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman's own health is seriously endangered by the pregnancy?"

```{r}
#| label: fig-endangeredhealth
#| echo: false
#| warning: false
#| fig-cap: Proportion of Yes and No Responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman's own health is seriously endangered by the pregnancy?"

# Endangered Health Responses
endangered_health_data <- cleaned_GSS_data[c('year', 'endangered_health')]

endangered_health_percentages <- endangered_health_data |>
  count(year, endangered_health) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

endangered_health_percentages$endangered_health <- factor(
  endangered_health_percentages$endangered_health, levels =
    responses)

ggplot(endangered_health_percentages, aes(
  x = as.factor(year),
  y = percentage, 
  fill = endangered_health)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Year",
       y = "Percentage",
       fill = "Responses") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), 
                     limits = c(0, 100)) + 
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal()
```

### Low Income

@fig-lowincome displays the proportion of Yes and No responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the family has a very low income and cannot afford any more children?"

```{r}
#| label: fig-lowincome
#| echo: false
#| warning: false
#| fig-cap: Proportion of Yes and No Responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the family has a very low income and cannot afford any more children?"

# Endangered Health Responses
low_income_percentages <- low_income_data |>
  count(year, low_income) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

low_income_percentages$low_income <- factor(
  low_income_percentages$low_income, levels =
    responses)

ggplot(low_income_percentages, aes(
  x = as.factor(year),
  y = percentage, 
  fill = low_income)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Year",
       y = "Percentage",
       fill = "Responses") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), 
                     limits = c(0, 100)) + 
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal()
```

### Abortion for Any Reason

@fig-anyreason displays the proportion of Yes and No responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason?"

```{r}
#| label: fig-anyreason
#| echo: false
#| warning: false
#| fig-cap: Proportion of Yes and No Responses in 2006 and 2016 to the question "Should it be possible for a pregnant woman to obtain a legal abortion if the woman wants it for any reason?"

# Endangered Health Responses
any_reason_percentages <- any_reason_data |>
  count(year, any_reason) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

any_reason_percentages$any_reason <- factor(
  any_reason_percentages$any_reason, levels =
    responses)

ggplot(any_reason_percentages, aes(
  x = as.factor(year),
  y = percentage, 
  fill = any_reason)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
       x = "Year",
       y = "Percentage",
       fill = "Responses") +
  scale_y_continuous(labels = scales::percent_format(scale = 1), 
                     limits = c(0, 100)) + 
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal()
```

# Results
## Respondent Demographics

@tbl-gender shows the count and percentage of male and female respondents for 2006 and 2016. Responses were filtered to be strictly male or female. For both years, the percentage of female participants were higher than the male participants, as the female participant percentages were always above 50% while the male participant percentages stayed close to the mid 40% range. 

```{r}
#| label: tbl-gender
#| echo: false
#| warning: false
#| tbl-cap: Respondent Gender Count and Percentage for 2006 and 2016

gender_demographic_data <- cleaned_GSS_data |>
  group_by(year, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

kable(gender_demographic_data, 
      col.names = c("Year", "Sex", "Count", "Percentage"),
      booktabs = TRUE,
      linesep = ""
      )
```

@tbl-agegroups displays the number of respondents among the different age groups for the 2006 and 2016 surveys. The classified age groups are ‘18-24’, ‘25-34’, ‘45-54’, ‘55-64’, and ‘65+’. The 18-24 age group had the least amount of participants each year, while the '35-44' age group had the highest proportion in 2006 and the '65+' age group had the highest proportion in 2016.

```{r}
#| label: tbl-agegroups
#| echo: false
#| warning: false
#| tbl-cap: Respondent Count of Participants in Age Groups in 2006 and 2016

age_group_2006 <- cleaned_GSS_data |>
  filter(year == 2006) |>
  mutate(age_group_2006 = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_group_2006) |>
  pivot_wider(names_from = age_group_2006, values_from = n)

age_group_2016 <- cleaned_GSS_data |>
  filter(year == 2016) |>
  mutate(age_group_2016 = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_group_2016) |>
  pivot_wider(names_from = age_group_2016, values_from = n)

# Put the Year column back
age_group_2006 <- mutate(age_group_2006, Year = 2006)
age_group_2016 <- mutate(age_group_2016, Year = 2016)

combined_age_group_table <- bind_rows(age_group_2006,
                                      age_group_2016)

combined_age_group_table <- combined_age_group_table |>
  select(Year, everything())

kable(combined_age_group_table,
      col.names = c("Year", "18-24", "25-34", "35-44", "45-54", "55-64", "65+"),
      booktabs = TRUE,
      linesep = ""
      )

```

@tbl-age shows that the age of the participants ranged from 18 to 89, with the average respondent being 47 years old in 2006 and 50 in 2016. Additionally, the mode age have a significant difference between the two years, with a mode value of 36 in 2006 and 52 in 2016. 

```{r}
#| label: tbl-age
#| echo: false
#| warning: false
#| tbl-cap: Total Respondent Mean, Median, Mode, Min, and Max Age by Year

age_2006_data <- data_2006 |>
  select('age') |>
   summarise(
    mean = round(mean(age, na.rm = TRUE)),
    mode = as.numeric(names(sort(table(age), decreasing = TRUE))[1]),
    min = min(age, na.rm = TRUE),
    max = round(max(age, na.rm = TRUE))
  )

age_2016_data <- data_2016 |>
  select('age') |>
   summarise(
    mean = round(mean(age, na.rm = TRUE)),
    mode = as.numeric(names(sort(table(age), decreasing = TRUE))[1]),
    min = min(age, na.rm = TRUE),
    max = round(max(age, na.rm = TRUE))
  )

combined_age_data <- bind_rows(age_2006_data, 
                               age_2016_data)
combined_age_data <- cbind(years, combined_age_data)


kable(combined_age_data, 
      col.names = c("Year",
                    "Mean", 
                    "Mode", 
                    "Min", 
                    "Max"),
      booktabs = TRUE,
      linesep = ""
      )

```

@tbl-religion displays the proportion of participants of each religion in the survey in 2006 and 2016. The religion with the highest proportion of responses for both years was "Protestant" with 52.46% in 2006 and 47.70% in 2016. The religion with the lowest proportion of responses for 2006 was "Other eastern religions" with a value of 0.11%, and there was an equal amount of respondents for "Other eastern religions" and "Inter-nondenominational" with a value of 0.24% in 2016.

@fig-allreligions graphically illustrates the proportions of the religious beliefs of the participants.

```{r}
#| label: tbl-religion
#| echo: false
#| warning: false
#| tbl-cap: Respondent Religion Count and Percentage for 2006 

religion2006_data <- data_2006 |>
  group_by(religion) |>
  summarise(count = n()) |>
  mutate(
    religion = factor(religion, levels = religion_order),
    percentage = round((count / sum(count)) * 100, 2)
  )
religion2016_data <- data_2016 |>
  group_by(religion) |>
  summarise(count = n()) |>
  mutate(
    religion = factor(religion, levels = religion_order),
    percentage = round((count / sum(count)) * 100, 2)
  )

religion2006_data_2 <- religion2006_data |>
  rename(count_2006 = count, percentage_2006 = percentage)

# Rename count and percentage columns in religion2016_cleaned_data
religion2016_data_2 <- religion2016_data |>
  rename(count_2016 = count, percentage_2016 = percentage)

# Combine the datasets using left_join
combined_data <- left_join(religion2006_data_2, 
                           religion2016_data_2,
                           by = "religion")

kable(combined_data,
      col.names = c("Religion",
      "2006 Count",
      "2006 Percentages",
      "2016 Count",
      "2016 Percentages"), 
      booktabs = TRUE,
      linesep = "")
```

```{r}
#| label: fig-allreligions
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Number of Respondents by Religion

religion_data <- cleaned_GSS_data |>
  count(religion) |>
  group_by(religion)

religion_data$religion <- factor(
  religion_data$religion, levels =
    religion_order)

ggplot(religion_data, aes(x = religion, y = n)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(x = "Religion", y = "Number of Respondents") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```
## Overall Trends

@tbl-avgresponses summarizes the average of the responses for each prompt, where 1 presents a "yes" response and 2 represents a "no" response. Each reason shows a decrease in the average from 2006 to 2016, showing that more participants are responding with "yes" in the later year. The most significant change was in the "Any Reason" reasoning, with a decrease of 0.06 in the value. Otherwise, the deceases remained minimal in the other prompts, ranging from 0.005 to 0.026. The decreases in value could be attributed to the change in societal attitudes and increased awareness of the issue at hand. 

```{r}
#| label: tbl-avgresponses
#| tbl-cap: Average of Responses by Year Divided by Question
#| warning: false
#| echo: false

rape_avg <- filtered_GSS_data |>
  group_by(year) |>
  summarize(
    avg_rape = round(mean(abrape), 3))

endangered_health_avg <- filtered_GSS_data |>
  group_by(year) |>
  summarize(
    avg_endangered_health = round(mean(abhlth), 3))

low_income_avg <- filtered_GSS_data |>
  group_by(year) |>
  summarize(
    avg_low_income = round(mean(abpoor), 3))

any_reason_avg <- filtered_GSS_data |>
  group_by(year) |>
  summarize(
    avg_any_reason = round(mean(abany), 3))

combined_table <- cbind(
  rape_avg,
  endangered_health_avg[, -1],
  low_income_avg[, -1],
  any_reason_avg[, -1])

kable(combined_table,
      format = "markdown", 
      col.names = c("Year",
                    "Rape",
                    "Endangered Health",
                    "Low Income",
                    "Any Reason"),
      digits = 3,
      booktabs = TRUE,
      linesep = "")
```

@fig-allproportions2006 displays the proportion of "yes" and "no" responses among all the variables in 2006. The rape and endangered health, have a noticeably higher amount of respondents who answered yes, with both proportions being higher then 75%. However, for the other two reasons, low income and any reason, a lower proportion of respondents responded yes, as they were both less than 50%. The differences in proportions could convey the US population's additional sympathy for health reasons that are beyond the mother's control.

@fig-allproportions2016 displays the proportion of "yes" and "no" responses among all the variables in 2016, with the patterns being very similar to the 2006 data. Although minimal, there is an increase in "yes" responses for the low income and any reason variables, illustrating a possible shift in societal attitudes regarding the issue. 

```{r}
#| label: fig-allproportions2006
#| echo: false
#| message: false
#| warning: false
#| fig-cap: Proportion (%) of Yes and No Responses to All Variables in 2006

# Prepare data
variable_data_2006 <- data_2006[c('rape', 
                                'low_income',
                                'endangered_health',
                                'any_reason')]

# Rape
rape_data_2006 <- data_2006[c('year','rape')] 

rape_percentages <- rape_data_2006 |>
  count(year, rape) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

rape_percentages$rape <- factor(
  rape_percentages$rape, levels =
    responses)

rape_proportions <- 
  ggplot(rape_percentages, mapping = aes(x = " ", y = percentage, fill = rape)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Rape", y = "Proportion for each response (%)", fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Low Income
low_income_data_2006 <- data_2006[c('year','low_income')] 
low_income_percentages <- low_income_data_2006 |>
  count(year, low_income) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

low_income_percentages$low_income <- factor(
  low_income_percentages$low_income, levels =
    responses)

low_income_proportions <- 
  ggplot(low_income_percentages, mapping = aes(x = " ", y = percentage, fill = low_income)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Low Income", 
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Endangered Health
endangered_health_data_2006 <- data_2006[c('year','endangered_health')] 
endangered_health_percentages <- endangered_health_data_2006 |>
  count(year, endangered_health) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

endangered_health_percentages$endangered_health <- factor(
  endangered_health_percentages$endangered_health, levels =
    responses)

endangered_health_proportions <- 
  ggplot(endangered_health_percentages, mapping = aes(x = " ", 
                                                      y = percentage,
                                                      fill = endangered_health)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Endangered Health",
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Any Reason
any_reason_data_2006 <- data_2006[c('year','any_reason')] 
any_reason_percentages <- any_reason_data_2006 |>
  count(year, any_reason) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

any_reason_percentages$any_reason <- factor(
 any_reason_percentages$any_reason, levels =
    responses)

any_reason_proportions <-
  ggplot(any_reason_percentages , mapping = aes(x = " ",
                                                y = percentage, 
                                                fill = any_reason)) +
  geom_col() +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Any Reason",
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

any_reason_proportion <- ggplot(any_reason_percentages, aes(x = year, y = percentage, fill = any_reason)) +
  geom_col() +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Any Reason", y = "", fill = "Response") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

combined_graph_2006 <- rape_proportions + 
  low_income_proportions +
  endangered_health_proportions +
  any_reason_proportions +
  plot_layout(nrow = 1)

combined_graph_2006
```

```{r}
#| label: fig-allproportions2016
#| echo: false
#| message: false
#| warning: false
#| fig-cap: Proportion (%) of Yes and No Responses to All Variables in 2016
# Prepare data
variable_data_2016 <- data_2016[c('rape', 
                               'low_income',
                                'endangered_health',
                                'any_reason')]

# Rape
rape_data_2016 <- data_2016[c('year','rape')] 

rape_percentages <- rape_data_2016 |>
  count(year, rape) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

rape_percentages$rape <- factor(
  rape_percentages$rape, levels =
    responses)

rape_proportions <- 
  ggplot(rape_percentages, mapping = aes(x = " ", y = percentage, fill = rape)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Rape", y = "Proportion for each response (%)", fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Low Income
low_income_data_2016 <- data_2016[c('year','low_income')] 
low_income_percentages <- low_income_data_2016 |>
  count(year, low_income) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

low_income_percentages$low_income <- factor(
  low_income_percentages$low_income, levels =
    responses)

low_income_proportions <- 
  ggplot(low_income_percentages, mapping = aes(x = " ", y = percentage, fill = low_income)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Low Income", 
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Endangered Health
endangered_health_data_2016 <- data_2016[c('year','endangered_health')] 
endangered_health_percentages <- endangered_health_data_2016 |>
  count(year, endangered_health) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

endangered_health_percentages$endangered_health <- factor(
  endangered_health_percentages$endangered_health, levels =
    responses)

endangered_health_proportions <- 
  ggplot(endangered_health_percentages, mapping = aes(x = " ", 
                                                      y = percentage,
                                                      fill = endangered_health)) +
  geom_col(show.legend = FALSE) +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Endangered Health",
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

# Any Reason
any_reason_data_2016 <- data_2016[c('year','any_reason')] 
any_reason_percentages <- any_reason_data_2016 |>
  count(year, any_reason) |>
  group_by(year) |>
  mutate(percentage = n / sum(n) * 100) 

any_reason_percentages$any_reason <- factor(
 any_reason_percentages$any_reason, levels =
    responses)

any_reason_proportions <-
  ggplot(any_reason_percentages , mapping = aes(x = " ",
                                                y = percentage, 
                                                fill = any_reason)) +
  geom_col() +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Any Reason",
       y = " ",
       fill = "Response") +
  theme(axis.title.x = element_text(size = 8)) 

any_reason_proportion <- ggplot(any_reason_percentages, aes(x = year, y = percentage, fill = any_reason)) +
  geom_col() +
  scale_fill_manual(values = custom_yes_no_colours) +
  theme_minimal() +
  labs(x = "Any Reason", y = "", fill = "Response") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

combined_graph_2016 <- rape_proportions + 
  low_income_proportions +
  endangered_health_proportions +
  any_reason_proportions +
  plot_layout(nrow = 1)

combined_graph_2016
```


### Age

Majority of the respondents were within the age range of 24-60 years old, as shown in @fig-allages. The highest point was at age 58 with 81 respondents and the lowest point of the graph was age 88 with 7 respondents. Despite the younger population showing more of an involvement in the topic, @fig-allages shows that there is still a substantial portion of the older generation contributing their opinion, conveying the importance of the topic among all ages. 

```{r}
#| label: fig-allages
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Number of Respondents by Age

age_data <- cleaned_GSS_data |>
  count(age) |>
  group_by(age)

ggplot(age_data, mapping = aes(x = age, y = n)) +
  geom_bar(stat = "identity") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 25)) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 25)) +
  labs(x = "Age", y = "Number of Respondents") +
  theme_classic() 

```

More specifically, the proportion of respondents organized by age group are shown in @tbl-age_rape, @tbl-age_endangeredhealth, 
@tbl-age_lowincome, and @tbl-age_anyreason, for the abortion reasons of rape, endangered health, low income, and any reason, respectively. The 65+ age group consistently demonstrated the highest proportion in no responses for each abortion reason, whereas the age group for yes responses varied, but stayed close in range. This may be correlated to the fact that older generations often hold more traditional and conservative views, in which abortion is not favoured upon. The 35-44 age group showed the highest proportion in yes responses for the reasons of rape, low income, and any reason, while the 45-54 age group held the highest amount of responses for the endangered health reason. The 35-54 age group may have more synpathy and put more emphasis on socioeconomic factors, health, and individuality compared to the other age groups. 

```{r}
#| label: tbl-age_rape
#| tbl-cap: Rape Response Proportion (%) by Age
#| warning: false
#| echo: false
#| message: false
age_rape_data <- cleaned_GSS_data[c('age','year', 'rape')]

age_group_rape <- age_rape_data |>
  group_by(rape) |>
  mutate(age_rape_data = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_rape_data) |>
  pivot_wider(names_from = age_rape_data, values_from = n) |>
  rename("Response" = `rape`) 

age_group_rape_percentage <- age_group_rape |>
  mutate(`18-24` = round(`18-24` / total_participants$count * 100, 2),
         `25-34` = round(`25-34` / total_participants$count * 100, 2),
         `35-44` = round(`35-44` / total_participants$count * 100, 2),
         `45-54` = round(`45-54` / total_participants$count * 100, 2),
         `55-64` = round(`55-64` / total_participants$count * 100, 2),
         `65+` = round(`65+` / total_participants$count * 100, 2))

switched_rows <- rbind(age_group_rape_percentage[2, ], age_group_rape_percentage[1, ], age_group_rape_percentage[-c(1, 2), ])

kable(switched_rows, 
      booktabs = TRUE,
      linesep = "")

```
```{r}
#| label: tbl-age_endangeredhealth
#| tbl-cap: Endangered Health Response Proportion (%) by Age
#| warning: false
#| echo: false
#| message: false
age_EH_data <- cleaned_GSS_data[c('age','year', 'endangered_health')]

age_group_EH <- age_EH_data |>
  group_by(endangered_health) |>
  mutate(age_EH_data = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_EH_data) |>
  pivot_wider(names_from = age_EH_data, values_from = n) |>
  rename("Response" = `endangered_health`) 

age_group_EH_percentage <- age_group_EH |>
  mutate(`18-24` = round(`18-24` / total_participants$count * 100, 2),
         `25-34` = round(`25-34` / total_participants$count * 100, 2),
         `35-44` = round(`35-44` / total_participants$count * 100, 2),
         `45-54` = round(`45-54` / total_participants$count * 100, 2),
         `55-64` = round(`55-64` / total_participants$count * 100, 2),
         `65+` = round(`65+` / total_participants$count * 100, 2))

switched_rows <- rbind(age_group_EH_percentage[2, ], age_group_EH_percentage[1, ], age_group_EH_percentage[-c(1, 2), ])

kable(switched_rows, 
      booktabs = TRUE,
      linesep = "")

```
```{r}
#| label: tbl-age_lowincome
#| tbl-cap: Low Income Response Proportion (%) by Age
#| warning: false
#| echo: false
#| message: false
age_LI_data <- cleaned_GSS_data[c('age','year', 'low_income')]

age_group_LI <- age_LI_data |>
  group_by(low_income) |>
  mutate(age_LI_data = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_LI_data) |>
  pivot_wider(names_from = age_LI_data, values_from = n) |>
  rename("Response" = `low_income`) 

age_group_LI_percentage <- age_group_LI |>
  mutate(`18-24` = round(`18-24` / total_participants$count * 100, 2),
         `25-34` = round(`25-34` / total_participants$count * 100, 2),
         `35-44` = round(`35-44` / total_participants$count * 100, 2),
         `45-54` = round(`45-54` / total_participants$count * 100, 2),
         `55-64` = round(`55-64` / total_participants$count * 100, 2),
         `65+` = round(`65+` / total_participants$count * 100, 2))

switched_rows <- rbind(age_group_LI_percentage[2, ], age_group_LI_percentage[1, ], age_group_LI_percentage[-c(1, 2), ])

kable(switched_rows, 
      booktabs = TRUE,
      linesep = "")

```
```{r}
#| label: tbl-age_anyreason
#| tbl-cap: Any Reason Response Proportion (%) by Age
#| warning: false
#| echo: false
#| message: false
age_AR_data <- cleaned_GSS_data[c('age','year', 'any_reason')]

age_group_AR <- age_AR_data |>
  group_by(any_reason) |>
  mutate(age_AR_data = case_when(
    age >= 18 & age <= 24 ~ "18-24",
    age >= 25 & age <= 34 ~ "25-34",
    age >= 35 & age <= 44 ~ "35-44",
    age >= 45 & age <= 54 ~ "45-54",
    age >= 55 & age <= 64 ~ "55-64",
    age >= 65 ~ "65+"
  )) |>
  count(age_AR_data) |>
  pivot_wider(names_from = age_AR_data, values_from = n) |>
  rename("Response" = `any_reason`) 

age_group_AR_percentage <- age_group_AR |>
  mutate(`18-24` = round(`18-24` / total_participants$count * 100, 2),
         `25-34` = round(`25-34` / total_participants$count * 100, 2),
         `35-44` = round(`35-44` / total_participants$count * 100, 2),
         `45-54` = round(`45-54` / total_participants$count * 100, 2),
         `55-64` = round(`55-64` / total_participants$count * 100, 2),
         `65+` = round(`65+` / total_participants$count * 100, 2))

switched_rows <- rbind(age_group_AR_percentage[2, ], age_group_AR_percentage[1, ], age_group_AR_percentage[-c(1, 2), ])

kable(switched_rows, 
      booktabs = TRUE,
      linesep = "")

```

### Sex

The gender demographic along with the responses for each abortion reason are exhibited and filtered by the years 2006 and 2016. The percentages are grouped by the response type of the participant, and then further grouped into their gender.
For each table, females made up a higher percentage for both responses in both of the years. Females showed the most support for abortions for the reason of endangered health (@tbl-endangeredhealthdemographic) and the least amount of support for abortions for any reason (@tbl-anyreasondemographic) in 2006 and  low income (@tbl-lowincomedemographic) in 2016.
Males also showed the most support for abortions in the case where the mother's health is endangered and showed the least amount of support for abortions that are obtained for any reason. While both genders conveyed a lot of support for abortions when there is risk of endangered health, participants also showed support for abortions for victims who were impregnated via rape (@tbl-rapedemographic).

```{r}
#| label: tbl-rapedemographic
#| tbl-cap: Rape Response Count and Proportions (%) by Sex
#| warning: false
#| echo: false

# 2006 table
rape_demographic_2006 <- data_2006[c('rape','sex')] 

rape_demographic_2006 <- rape_demographic_2006 |>
  mutate(response = factor(rape, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# 2016 table
rape_demographic_2016 <- data_2016[c('rape','sex')] 

rape_demographic_2016 <- rape_demographic_2016 |>
  mutate(response = factor(rape, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# Combine tables
combined_R_table <- inner_join(rape_demographic_2006,
                               rape_demographic_2016, 
                               by = c("response", "sex"))
kable(combined_R_table,
      col.names = c("Response",
                    "Sex",
                    "Count",
                    "Percentage",
                    "Count",
                    "Percentage"),
      booktabs = TRUE,
      linesep = "") |>
  add_header_above(c( " " = 2, "2006" = 2, "2016" = 2), bold = TRUE)
```

```{r}
#| label: tbl-endangeredhealthdemographic
#| tbl-cap: Endangered Health Response Count and Proportions (%) by Sex
#| warning: false
#| echo: false

# 2006 table
endangered_health_demographic_2006 <-
  data_2006[c('endangered_health','sex')] 

endangered_health_demographic_2006 <-
  endangered_health_demographic_2006 |>
  mutate(response = factor(endangered_health, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# 2016 table
endangered_health_demographic_2016 <- 
  data_2016[c('endangered_health','sex')] 

endangered_health_demographic_2016 <-
  endangered_health_demographic_2016 |>
  mutate(response = factor(endangered_health, 
                           levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# Combine tables
combined_EH_table <- inner_join(endangered_health_demographic_2006,
                               endangered_health_demographic_2016, 
                               by = c("response", "sex"))
kable(combined_EH_table,
      col.names = c("Response",
                    "Sex",
                    "Count",
                    "Percentage",
                    "Count",
                    "Percentage"),
      booktabs = TRUE,
      linesep = "") |>
  add_header_above(c( " " = 2, "2006" = 2, "2016" = 2), bold = TRUE)
```

```{r}
#| label: tbl-lowincomedemographic
#| tbl-cap: Low Income Response Count and Proportions (%) by Sex
#| warning: false
#| echo: false

# 2006 table
low_income_demographic_2006 <-
  data_2006[c('low_income','sex')] 

low_income_demographic_2006 <- low_income_demographic_2006 |>
  mutate(response = factor(low_income, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# 2016 table
low_income_demographic_2016 <- data_2016[c('low_income','sex')] 

low_income_demographic_2016 <- low_income_demographic_2016 |>
  mutate(response = factor(low_income, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# Combine tables
combined_LI_table <- inner_join(low_income_demographic_2006,
                             low_income_demographic_2016, 
                             by = c("response", "sex"))
kable(combined_LI_table,
      col.names = c("Response",
                    "Sex",
                    "Count",
                    "Percentage",
                    "Count",
                    "Percentage"),
      booktabs = TRUE,
      linesep = "") |>
  add_header_above(c( " " = 2, "2006" = 2, "2016" = 2), bold = TRUE)
```

```{r}
#| label: tbl-anyreasondemographic
#| tbl-cap: Any Reason Response Count and Proportions (%) by Sex
#| warning: false
#| echo: false

# 2006 table
any_reason_demographic_2006 <- data_2006[c('any_reason','sex')] 

any_reason_demographic_2006 <- any_reason_demographic_2006 |>
  mutate(response = factor(any_reason, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# 2016 table
any_reason_demographic_2016 <- data_2016[c('any_reason','sex')] 

any_reason_demographic_2016 <- any_reason_demographic_2016 |>
  mutate(response = factor(any_reason, levels = responses)) |>
  group_by(response, sex) |>
  summarise(count = n()) |>
  mutate(percentage = round((count / sum(count)) * 100, 2))

# Combine tables
combined_AR_table <- inner_join(any_reason_demographic_2006,
                               any_reason_demographic_2016, 
                               by = c("response", "sex"))
kable(combined_AR_table,
      col.names = c("Response",
                    "Sex",
                    "Count",
                    "Percentage",
                    "Count",
                    "Percentage"),
      booktabs = TRUE,
      linesep = "") |>
  add_header_above(c( " " = 2, "2006" = 2, "2016" = 2), bold = TRUE)
```

### Religious Views

Narrowing down the religions, @fig-religionprop displays the proportion of yes and no responses specifically for the Protestant and Non-Religious Respondents. The Protestant group was chosen since it was the religious group with the highest proportion of respondents, and the Non-Religious group was chosen to compare how religion may affect one's perspective on the issue. 
As we can see from @fig-religionprop, a higher proportion of protestant answered "no" consistently for each reason for abortion. These responses may be due to moral teachings and inteprations of life within the Christian religion, prompting more people to believe that abortion is taking away a life. Both Non-Religious and Protestant respondents showed the least support for rape cases or low income for reasons to obtain a legal abortion. Although the fraction of Protestants choosing "no" is higher, the graph illustrates how there is a shared perspective among the two groups regarding perspectives on abortion in cases involving rape or financial hardship. 

```{r}
#| label: fig-religionprop
#| fig-cap: Proportion of Responses from Protestant and Non-Religious Participants 
#| warning: false
#| echo: false

prot_none <- cleaned_GSS_data |>
  filter(religion %in% c("Protestant", "None")) |>
  select(religion, rape, low_income, endangered_health, any_reason)

prot_none_data <- prot_none |>
  pivot_longer(cols = c(rape, low_income, endangered_health, any_reason),
               names_to = "reason",
               values_to = "response")

ggplot(prot_none_data, aes(x = reason, fill = response)) +
  geom_bar(position = "fill", stat = "count") +
  facet_wrap(~ religion, ncol = 1) +
  labs(x = "Reasons for Abortion",
       y = "Proportion",
       fill = "Response") +
  scale_x_discrete(labels = reasons) +
  scale_fill_manual(values = custom_yes_no_colours_reversed) +
  guides(fill = guide_legend(reverse = TRUE)) +
  theme_minimal() +
  theme(legend.position = "right",
        axis.text.x = element_text(size = 8), 
        axis.title = element_text(size = 10),   
        strip.text = element_text(size = 14)) 
```

# Discussion

## Generational Differences

## Sex

## Religious Views

## Economic Status

## Temporal Movements and Law Changes

-   will talk about the different law changes and movements occurring before, during, and after the specific years

### 2006

### 2016

\newpage

\appendix

# Appendix {.unnumbered}

asdf

\newpage

# References
